---
title: "Simulation of dataset of price-responsive buildings"
author: "Javier Saez Gallego"
date: "July 26, 2016"
output: html_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Peparing the dataset

Here we simulate 100 buildings that are equipped with a home automation system, that receoved a price of electricity and scheules the consumption such that the total cost is minimized. The heating dynamics of each building is described by a state-space model that consists of three states: indoor air temperature, floor temperature, and temperature of the water inside a tank connected to a heat pump. The only input is the electricity consumption.


$$\underset{ \boldsymbol y, \boldsymbol x ,\boldsymbol v }{\text{Minimize}}  \sum_{t=1}^T p_t  x_t + \rho v_t  	$$
$$ \boldsymbol y_{t} = \boldsymbol  A \boldsymbol y_{t-1} + \boldsymbol B x_{t-1} + \boldsymbol E  \boldsymbol z_{t-1} 	\   \forall t 	$$
$$  0 \leq x_t \leq x^{max} 	 \forall t 	$$
$$  y_{t,min}^r \leq y_t^r + v_t   	 \forall t 	$$
$$ y_{t,max}^r \geq y_t^r - v_t   	 \forall t $$
$$ v_t \geq 0    	 \forall t 	$$

A total of 100 buildings are simulated by randomly perturbing the heat-transfer coefficients. Modifying slightly these coefficients allows us to simulate the behavior of buildings with different structural characteristics. The perturbations are randomly drawn from a uniform distribution centered around zero with a variance equal to 1/50 the magnitude of the corresponding coefficient. The magnitude of the perturbations is chosen high enough so that different building structures are modeled, but not too high so that the state-space system becomes unstable. The magnitude of the perturbations has been chosen by trial-and-error, and its effectiveness is proven to be useful as explained in the remaining of this section and in the case study

Together with the simulated houses, we algo put together some calendar variables that will help explain the variabiliy of the parameters.

## Initialization

First, we define the working paths of R. Also, we load GAMS, the auxiliary functions, and the [R_to_GDX](https://github.com/jsga/R_to_GAMS) functions. We create a working dorectory to allocate the GAMS auxiliary files.

```{r,eval=FALSE}
wpath = '/home/jsga/inverse_opt/'
gams_file = "/home/jsga/GAMS23.7/gams"
library('gdxrrw')
igdx('/home/jsga/GAMS24.2.1/gams24.2_linux_x64_64_sfx')

newDir = paste(wpath, 'Working_dir',sep="")
dir.create(newDir)
setwd(newDir)

source(paste(wpath,"AuxiliaryFunctions.R",sep=""))
source(paste(wpath,"R_to_GDX.r",sep=""))
```



## Input parameters

Define the parameters that define the "average" house (taken from _Zugno et al._ "A bilevel model for electricity retailers' participation in a demand response market environment").

```{r,eval=FALSE}
T=24
D=28*3
set.seed(100)
price = rep(c(32.02,31.08,30.62,30.19,30.36,29.95,36.41,38.09,38.82,38.23,37.80,37.37,37.12,37.22,37.07,36.95,37.58,37.68,36.98,36.82,35.86,32.38,32.09,29.12),D) + rnorm(T*D,0,3)
pv = 1000
a11 = 0.4103  # Tr(t-1) -> Tr
a12 = 0.5586  # Tf(t-1) -> Tr
a13 = 0.0028  # Tw(t-1) -> Tr

a21 = 0.1092  # Tr(t-1) ->Tf
a22 = 0.87    # Tf(t-1) ->Tf
a23 = 0.0078  # Tw(t-1) ->Tf

a31 = 0.0022  # Tr(t-1) -> Tw
a32 = 0.031   # Tf(t-1) -> Tw 
a33 = 0.975   # Tw(t-1) -> Tw

b1 = 0.0044   # Wc(t-1) -> Tr
b2 = 0.0173   # Wc(t-1) -> Tf
b3 =  4.2332  # Wc(t-1) -> Tw

d1 = 0.01    # S(t-1) -> Tr
d2 = 0.001  # S(t-1) -> Tf

e1 = 0.0029   # Ta(t-1) -> Tr
e2 = 0        # Ta(t-1) -> Tf
e3 = 0        # Ta(t-1) -> Tw

maxHP = 0.33 # Kwh
minHP = 0
rampUpHP = maxHP/1
rampDownHP = -maxHP/1
S = rep(c( 0,0,0,0,0,0,50*sin( 2*pi*(1:12)/24) + rgeom(12,0.1),0,0,0,0,0,0),times=D)*rlnorm(T*D, meanlog = 0, sdlog = 0.4)
Ta = 5 + 10*cos( 2*pi*12:(T*D+11)/24) + rnorm(T*D,0,5)
```

It is important to note the comfort bands, defined below. For non-responsive houses we can set the bands to be equal:

```{r,eval=FALSE}
meanComfort = rep(c(20,20,20,20,20,20,20,21,21,21,21,21,21,21,21,21,21,21,21,21,21,20,20,20),times=D)
maxComfort = meanComfort+1
minComfort = meanComfort-1
```


## Simulation in a loop

We simulate different houses by randomly perturbing the heat-transfer coefficients. We generate 3 times more houses than we need and then discard the ones with highest "discomfort". The reason is that, for some of the simulated state-space model, the system is unstable. In this way we just keep the "good" ones.

The function `heat_pump_sim()` is defined in `AuxiliaryFunctions.R`.
```{r,eval=FALSE}
H = 100# Number of houses
WC_all_l = matrix(NA,ncol=0,nrow=T*D)
Tr_all_l = matrix(NA,ncol=0,nrow=T*D)
Tf_all_l = matrix(NA,ncol=0,nrow=T*D)
Tw_all_l = matrix(NA,ncol=0,nrow=T*D)
D_all_l = c()
param_pool = list()
stop=F # Stop bool
i=1 # Count number of saved simulation
l = 1 # Count rejected simulations

while(stop==F){
  # Save the parameters of the population for later use
  K=50
  param_pool$a11[i] = a11+runif(1,-a11/K,a11/K)
  param_pool$a12[i] = a12+runif(1,-a12/K,a12/K)
  param_pool$a13[i] = a13+runif(1,-a13/K,a13/K)
  param_pool$a21[i] = a21+runif(1,-a21/K,a21/K)
  param_pool$a22[i] = a22+runif(1,-a22/K,a22/K)
  param_pool$a23[i] = a23+runif(1,-a23/K,a23/K)
  param_pool$a31[i] = a31+runif(1,-a31/K,a31/K)
  param_pool$a32[i] = a32+runif(1,-a32/K,a32/K)
  param_pool$a33[i] = a33+runif(1,-a33/K,a33/K)
  
  param_pool$b1[i] = b1
  param_pool$b2[i] = b2
  param_pool$b3[i] = b3
  
  param_pool$d1[i] = d1#+runif(1,-d1/K,d1/K)
  param_pool$d2[i] = d2#+runif(1,-d2/K,d2/K)
  
  param_pool$e1[i] = e1 
  param_pool$e2[i] = e2
  param_pool$e3[i] = e3
  
  # Perform the optimization
  HP_sol = heat_pump_sim(price,Ta,S,pv, # Price, temperature and solar are the same for all
                         param_pool$a11[i],param_pool$a12[i],param_pool$a13[i],
                         param_pool$a21[i],param_pool$a22[i],param_pool$a23[i],
                         param_pool$a31[i],param_pool$a32[i],param_pool$a33[i],
                         param_pool$b1[i],param_pool$b2[i],param_pool$b3[i],
                         param_pool$d1[i],param_pool$d2[i],
                         param_pool$e1[i],param_pool$e2[i],param_pool$e3[i],
                         maxHP,minHP,rampUpHP,rampDownHP,minComfort,maxComfort,
                         folder_file = paste(wpath,"GAMS_models/",sep=""),N=T*D)
  if(HP_sol$status != 1){
    message(paste( 'Status sol not optimal',HP_sol$status))
    Sys.sleep('5')
  }
  
  WC_all_l = cbind(WC_all_l,HP_sol$Wc)
  Tr_all_l = cbind(Tr_all_l,HP_sol$Tr)
  Tf_all_l = cbind(Tf_all_l,HP_sol$Tf)
  Tw_all_l = cbind(Tw_all_l,HP_sol$Tw)
  D_all_l = c(D_all_l,sum(HP_sol$D))
  # Increase iteration of saved
  i = i+1
  # Post messahe
  if(i%%50 == 0) message(paste("sim nr",i))
  
  # Check for end of loop
  if(i > 3*H){
    stop=TRUE
    system('notify-send Finished loop')
  }
  
}

# Keep the simulated values with lowest discomfort
i_keep = order(D_all_l)[1:H]
WC_all = WC_all_l[,i_keep]
Tr_all = Tr_all_l[,i_keep]
Tf_all = Tf_all_l[,i_keep]
Tw_all = Tw_all_l[,i_keep]
D_all = D_all_l[i_keep]
load_all = rowSums(WC_all)
```



## Generate regressors

Here we create some time and calendar variables to support the estimation process.

```{r,eval=FALSE}
# Create time indicator
Time = rep(0:23,D)

# Create dummy variables for time
## Weekly fourier
fourier = function(t,terms,period,names_col=c("S","C")){
  n <- length(t)
  X <- matrix(,nrow=n,ncol=2*terms)
  for(i in 1:terms){
    X[,2*i-1] <- sin(2*pi*i*t/period)
    X[,2*i] <- cos(2*pi*i*t/period)
  }
  colnames(X) <- paste(names_col,rep(1:terms,rep(2,terms)),sep="")
  return(X)
}

N = length(Time)
data_hour = scale(data.frame(fourier(1:N,16,96,c("Daily_S","Daily_C"))))

# Add binary indicator of hour
data_hour_bin = matrix(,ncol=24,nrow=N)
for(i in 1:24) data_hour_bin[,i] = (as.numeric(Time == i-1))

data_hour_bin = data.frame(data_hour_bin)
colnames(data_hour_bin) = paste('H',0:23,sep=".")
```


Also, we lag the simulated load and the price. Past values of the load are used to explain the current load.

```{r,eval=FALSE}
data_aux = data.frame( Time = Time,
                       price = price,
                       load = load_all,
                       S = S,
                       Ta = Ta,
                       data_hour,
                       data_hour_bin)

# Load lag function
source(paste(wpath,'lagmatrix.R',sep=""))

# Lag the load
n_lag = 24
load_lag  = lagmatrix(load_all,n_lag,prenames="load")
# Merge with dataset- Remove n_lag first observations
data_lag = cbind(data_aux[-(1:n_lag),],load_lag)

# Lag the price
price_lag  = lagmatrix(data_aux$price,n_lag,prenames="price")
# Merge with dataset- Remove n_lag first observations
data_lag = cbind(data_lag,price_lag) # the first 300 obs are already removed

# Replace
data = data_lag
```

## Save the dataset and finish

Lastly: save the dataset in a nice format.

```{r,eval=FALSE}
data_sim = list(
  WC = WC_all,
  Tr = Tr_all,
  Tf = Tf_all,
  Tw = Tw_all
)

data_list =      list(data = data,
                      data_sim = data_sim,
                      D = D,
                      maxHP = maxHP,
                      minHP = minHP,
                      rampUpHP = rampUpHP,
                      rampDownHP = rampDownHP,
                      maxComfort = maxComfort,
                      minComfort = minComfort,
                      H = H,
                      K= K)
str(data_list)
save(data_list ,file=paste(wpath,"data_2flex')
```

